---
title: "Presentation"
author: "Chunjingwen Cui and MING HAO LEE "
date: "2023-03-06"
output: html_document
---
This is a R project that analysis a data set including different data analysis. 




# Data Set
Life Expectancy (WHO)
The Global Health Observatory (GHO) data repository under World Health Organization (WHO) keeps track of the health status as well as many other related factors for all countries The data-sets are made available to public for the purpose of health data analysis. The data-set related to life expectancy, health factors for 193 countries has been collected from the same WHO data repository website and its corresponding economic data was collected from United Nation website. 


The reason we want ot use this dataset is about life expectancy, including data related to life expectancy in different years in different countries. This data set has a large amount of data, including 22 columns and 2938 rows, which meets the requirements of this project. In addition, many factors affect life expectancy, and this dataset lists immune-related, mortality, economic, and social factors. By analyzing these data, we can understand how these factors affect humanlifeexpectancy.We can also make recommendations to the government to improve the population's life expectancy based on the data analysis results.


The available variables are:


Country: 193 countries	


Year: 2000-2015


Status: Developed or Developing status


Life expectancy: Life Expectancy in age	


Adult Mortality: Adult Mortality Rates of both sexes (probability of dying between 15 and 60 years per 1000 population)


infant deaths: Number of Infant Deaths per 1000 population


Alcohol: Alcohol, recorded per capita (15+) consumption (in litres of pure alcohol)


percentage expenditure: Expenditure on health as a percentage of Gross Domestic Product per capita(%)	


Hepatitis B: Hepatitis B (HepB) immunization coverage among 1-year-old (%)


Measles: Measles - number of reported cases per 1000 population 


BMI: Average Body Mass Index of entire population


under-five deaths: Number of under-five deaths per 1000 population


Polio: Polio (Pol3) immunization coverage among 1-year-old (%)


Total expenditure: General government expenditure on health as a percentage of total government expenditure (%)


Diphtheria: Diphtheria tetanus toxoid and pertussis (DTP3) immunization coverage among 1-year-old (%)


HIV/AIDS: Deaths per 1 000 live births HIV/AIDS (0-4 years)


GDP: Gross Domestic Product per capital (in USD)	


Population: Population of the country 


thinness  1-19 years: Prevalence of thinness among children and adolescents for Age 10 to 19 (%)


thinness 5-9 years: Prevalence of thinness among children for Age 5 to 9(%)


Income composition of resources: Human Development Index in terms of income composition of resources (index ranging from 0 to 1)


Schooling: Number of years of Schooling(years)


 
# Start Anlysis
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We'll start by reading in the data:
```{r}
library(tidyverse)
setwd("/Users/minghaolee/Desktop/UOP_Class/COMP_293/Project")
lifeExpectancy<- read_csv("Life Expectancy Data.csv")
lifeExpectancyP1 <- lifeExpectancy[,c("Status","Life expectancy", "Alcohol","Schooling")]
```


## Exploratory data analysis
### Descriptive statistics
Let's start by finding some summaries of our variables.

```{r}
str(lifeExpectancy)
summary(lifeExpectancy)
```
Based on this output, we have an interesting phenomenon that the Min of min of some variables are 0, maybe because the data lost. 





Let's try to see how many different developing and developed countries within this dataset
```{r}
lifeExpectancy %>%
  group_by(Status) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```

Looks like there are 2426 developing countries more against 512 developed countries. 



Let's then see which countries has the Longest life expectancy! 
```{r}
lifeExpectancy %>%
  group_by(Country) %>%
  summarize(avg_LE = mean(`Life expectancy`)) %>%
  arrange(desc(avg_LE))
```
Japan is win! about 82.53 is the highest life expectancy. 

Then let's look at which countries has the most high average of alcohol consumption. 
```{r}
lifeExpectancy %>%
  group_by(Country) %>%
  summarize(avg_Al = mean(Alcohol)) %>%
  arrange(desc(avg_Al))
```
Base on the result, United Kingdom of Great Britain and Northern Ireland is the winner which data is around 11.13. 

Finally, let's take a look at the difference average schooling between developed countries and developing countries. 
```{r}
lifeExpectancy %>%
  group_by(Status) %>%
  drop_na(Schooling)%>%
  summarize(avg_Sc = mean(Schooling)) %>%
  arrange(desc(avg_Sc))
```
Base on the result, the developed is 15.8 which is more then 11.2 of developing countries. 


### Exploratory visualizations
Now let's look at some visualizations of our data. Let's look at the relationships between our quantitative variables:
```{r}
lifeExpectancy %>%
  ggplot(aes(Status, `Life expectancy`)) +
  geom_boxplot()
```

Based on the output, we could find out that longer schooling, longer life expectation on both Developed or Developing countries. 
Whereas there do have some outliers of schooling. 



Now let's look at some visualizations of our data. Let's look at the relationships between the three quantitative variables that we focusing on:
```{r}
lifeExpectancy %>%
  select(`Life expectancy`, Schooling, Alcohol) %>%
  pairs()
```

Our analysis focuses on three quantitative variables - life expectancy, schooling, and alcohol. Looks like there are some interesting relationships here. 
Let's take a look about there correlation 
```{r}
lifeExpectancy %>%
  select(`Life expectancy`, Schooling, Alcohol, GDP) %>%
  drop_na() %>%
  cor()
```
In this result we could see their revealed positive correlations between every two variables, lets see how about the spareman correlation
```{r}
lifeExpectancy %>%
  select(`Life expectancy`, Schooling, Alcohol, GDP) %>%
  drop_na() %>%
  cor(method = 'spearman')
```
We can see every variables turns more strong positive relationship, particularly between life expectancy and schooling. 


The correlation between life expectancy and schooling is strong and appears to be linear. As a result, we plan to conduct a linear regression analysis using these two variables.
```{r}
cor.test(lifeExpectancy$`Life expectancy`, lifeExpectancy$Schooling)
```

Lets see how about in Spearman correlation
```{r}
cor.test(lifeExpectancy$`Life expectancy`, lifeExpectancy$Schooling, method = "spearman")
```

```{r}
lifeExpectancy %>%
  filter(Schooling > 0) %>%
  drop_na() %>%
  ggplot(aes(Schooling, `Life expectancy`)) + 
  geom_point(aes(color = Status)) + 
  geom_smooth(method = lm)
```


Base on the output, the result is pretty obvious. a clear strong linear regression. 


The null hypothesis that we state: 


The slope of the regression line for schooling and life expectancy is equal to 0.


The p-value is significantly less than 0.05, so we can reject null hypothesis; and the r-square is 0.61, indicating that approximately 61% of the variation is accounted for by this regression model. Besides, the Pearson and Spearman correlation coefficients are 0.75 and 0.81, respectively.



Then, lets do a t-test to investigate whether there is a difference in life expectancy between developed and developing countries.
```{r}
t.test(`Life expectancy` ~ Status, data = lifeExpectancy)
lifeExpectancy %>%
  ggplot(aes(Status, `Life expectancy`)) +
  geom_boxplot() 
```

The null hypothesis that we state: 

that the difference in average life expectancy between the two groups is zero.

Base on the t-test, founding that the resulting p-value was much less than 0.05. Therefore, we can reject the null hypothesis, indicating that the mean life expectancy differs significantly between developed and developing countries.



Moreover,based on our earlier analysis, we have learned that longer schooling is associated with longer life expectancy, and that developed countries tend to have longer average schooling years. To explore whether individuals with higher education or prolonged education have equal life expectancies in developed and developing countries, we focused on data with 15 or more years of schooling. 
```{r}
subset <- lifeExpectancy %>%
  filter(Schooling > 0)
summary(lm(`Life expectancy` ~ Schooling, data = subset))

# t-test
t.test(`Life expectancy` ~ Status, data = lifeExpectancy)

schooling15 <- filter(lifeExpectancy, Schooling >= 15)
t.test(Schooling ~ Status, data = schooling15)

schooling15 %>%
  ggplot(aes(Status, Schooling)) +
  geom_boxplot()

```

Our t-test yielded a p-value significantly below 0.05, allowing us to reject the null hypothesis that there is no difference in the average life expectancy between developed and developing countries for such individuals. Our conclusion is also supported by the boxplot.


### Conclusion
The conclusion about this dataset base on our analysis research is that longer education, longer life expectancy. And developed countries have a longer life expectancy even considering only those with long-term education. Besides, For alcohol consumption, the consumption in developed countries is greater than that in developing countries, and the longer the education time, the higher the consumption. Whereas we still have a lot of other factors needs we to analyzing more, we will did the further research in the future. 


